Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-05-14 doc/report/msc-report.tex
problem-oriented abstract.

This thesis describes a data store for multi-dimensional time-series-based
data. The data is aggregated over many units of measurements and across
multiple dimensions. One of these dimensions is always time.

We present a data store called Zohmg. It is built on top of HBase and uses
Hadoop to run import jobs in parallel. Zohmg stores and serves aggregates of
measurements. The measurements are extracted taken from log files or similar
sources with user-supplied code.

The data store is optimized for speed of data retrieval: one of the design
goals was to serve data at mouse-click rate and promote real-time data
exploration. As such, Zohmg stores only the summarized data, not the atomic
measurements.

Similar data stores exist: they generally use a relational database system as
their underlying data store. The novelty of our approach is to model OLAP
cubes on top of a bigtable-like data store.

>>>>

Making sense of data at a massive scale is one of the biggest challenges we face
at Last.fm. Interpreting patterns in user behaviour becomes a challenge when
millions of users interact in billions of combinations; the data sets must be
analyzed, summarized and presented visually in order to convey their meaning.

This thesis describes a data store for multi-dimensional time-series-based data.
Measurements are summarized across multiple dimensions. The data store is
optimized for speed of data retrieval: one of the design goals is to serve data
at mouse-click rate to promote real-time data exploration.

Similar data stores exist; they generally use relational database systems as
their backing database. The novelty of our approach is to model multidimensional
data cubes on top of a column-oriented database, reaping the scalability
benefits of such databases.
